### 2020-08-14 TCS_v3/00-abstract.tex
nuovi file TCS v3

In the area of Natural Computing, Reaction Systems (RSs) are
a qualitative abstraction inspired by the 
functioning of living cells, suitable to model the main
mechanisms of biochemical reactions.
%This model has already been applied
%and extended successfully to various areas of research. 
RSs interact with a context, and
pose challenges for modularity, compositionality, extendibility and behavioural equivalence. 
In this paper we define a modular encoding of RSs as processes
in the chained Core Network Algebra (cCNA), which is a new variant  of the  {\tt link}-calculus.
The encoding represents the behaviour of each entity separately and
preserves faithfully their features, and we prove its correctness and completeness.
%we consider a variant of the {\tt link}-calculus, which allows to model 
%multiparty interaction in concurrent systems, and show that we can
%represent reaction systems as processes 
%in a modular way, by representing the behaviour of each entity and
%preserving faithfully their features.
%We show the correctness and completeness 
%of our embedding.
%We illustrate our framework by 
%{\color{red} We show} the embedding of 
%a few examples expressing computer science
%and biological applications. 
Our encoding provides a Labelled Transition System (LTS) semantics for RSs. Based on the LTS semantics, we adapt  the classical notion of  bisimulation to define a novel equivalence, called bio-similarity, for studying properties of RSs.
%On top of the LTS semantics for reaction systems provided by our framework, 
%we then adapt the classical notion (in concurrency) of bisimulation
%to make it more suitable for studying properties of reaction systems.
In particular, we define a new assertion language based on regular expressions, 
which allows to specify the properties of interest, and use it to extend Hennessy-Milner logic
to our setting.
% framework. 
We prove that our bio-similarity relation and the logical equivalence, that are defined parametrically on some assertion of interest, coincide.
%Interestingly, the novel notion of bisimilarity and logical equivalence that are defined parametrically on some assertion of interest are proved to coincide, like in the classical case.
Finally, 
we claim that
our encoding 
%can 
contributes to increase the expressiveness
of RSs, by exploiting the interaction among 
different RSs.

>>>>

In the area of Natural Computing, Reaction Systems (RSs) are
a qualitative abstraction inspired by the 
functioning of living cells, suitable to model the main
mechanisms of biochemical reactions.
%This model has already been applied
%and extended successfully to various areas of research. 
RSs interact with a context, and
pose challenges for modularity, compositionality, extendibility and behavioural equivalence. 
In this paper we define a modular encoding of RSs as processes
in the chained Core Network Algebra (cCNA), which is a new variant  of the  {\tt link}-calculus.
The encoding represents the behaviour of each entity separately and
preserves faithfully their features, and we prove its correctness and completeness.
%we consider a variant of the {\tt link}-calculus, which allows to model 
%multiparty interaction in concurrent systems, and show that we can
%represent reaction systems as processes 
%in a modular way, by representing the behaviour of each entity and
%preserving faithfully their features.
%We show the correctness and completeness 
%of our embedding.
%We illustrate our framework by 
%{\color{red} We show} the embedding of 
%a few examples expressing computer science
%and biological applications. 
Our encoding provides a Labelled Transition System (LTS) semantics for RSs. Based on the LTS semantics, we adapt  the classical notion of  bisimulation to define a novel equivalence, called bio-similarity, for studying properties of RSs.
%On top of the LTS semantics for reaction systems provided by our framework, 
%we then adapt the classical notion (in concurrency) of bisimulation
%to make it more suitable for studying properties of reaction systems.
In particular, we define a new assertion language based on regular expressions, 
which allows us to specify the properties of interest, and use it to extend Hennessy-Milner logic
to our setting.
% framework. 
We prove that our bio-similarity relation and the logical equivalence, that are defined parametrically on some assertion of interest, coincide.
%Interestingly, the novel notion of bisimilarity and logical equivalence that are defined parametrically on some assertion of interest are proved to coincide, like in the classical case.
Finally, 
we claim that
our encoding 
%can 
contributes to increase the expressiveness
of RSs, by exploiting the interaction among 
different RSs.

### 2020-08-14 TCS_v3/01-introduction.tex
nuovi file TCS v3

Our main contributions are as follows:
\begin{itemize}
\item 
\begin{color}{blue}
RSs are encoded in a modular way, in the sense that all constituents of a RS, namely entities, reactions and the context, are seen as interacting  c\CNA \ processes; in principle, this allows to study the behaviour of any constituent in isolation, contrary to what happens in the basic RS framework, where the evolution is possible only when all constituents are given;
\end{color}

@@

Moreover, we sketch how the encoding can be used to enhance the expressivity of RSs along different dimensions:
\begin{itemize}
\item we sketch how one can express the behaviour of entity mutation,
%change the behaviour of  each single entity $s$  in a
%way that a mutation $s'$ can be expressed; 
in such a way that the mutated version of the entity
$s$ can  take part to only a subset of rules requiring entity $s$;
\item we sketch how with a little coding effort, our 
%encoding 
embedding
allows two RSs to
communicate; i.e. we can model scenarios where a subset of those entities that the context  can
provide, are instead provided by a second RS.
\end{itemize}

>>>>

Our main contributions are as follows:
\begin{itemize}
\item 
RSs are encoded in a modular way, in the sense that all constituents of a RS, namely entities, reactions and the context, are seen as interacting  c\CNA \ processes; in principle, this allows to study the behaviour of any constituent in isolation, contrary to what happens in the basic RS framework, where the evolution is possible only when all constituents are given;

@@

Moreover, we sketch how the encoding can be used to enhance the expressivity of RSs along different dimensions:
\begin{itemize}
\item we sketch how one can express the behaviour of entity mutation,
%change the behaviour of  each single entity $s$  in a
%way that a mutation $s'$ can be expressed; 
in such a way that the mutated version of the entity
$s$ can  take part to only a subset of reactions requiring entity $s$;
\item we sketch how with a little coding effort, our 
%encoding 
embedding
allows two RSs to
communicate; i.e. we can model scenarios where a subset of those entities that the context  can
provide, are instead provided by a second RS.
\end{itemize}

### 2020-08-14 TCS_v3/02-ReactionSystem.tex
nuovi file TCS v3

\begin{definition}[Reaction]\label{def:reaction}
Let $S$ be a set of entities.
A reaction over $S$ is a triplet $a = (R,I,P)$, where $R, I, P$ are finite,  
non empty subsets of $S$ and
$R \cap I = \emptyset$. 
%If $S$ is a set such that  $R, I, P \subseteq S$, 
%then $a$ is a reaction in $S$.
\end{definition}

@@

\begin{definition}[Reaction Result]
Given a set of entities $S$, let $W\subseteq S$ be a finite subset of  entities.
\begin{enumerate}
\item Let $a$ be a reaction over $S$. Then $a$ is enabled by the entities in the set $W$, denoted by $\mathit{en}_a(W)$, 
if $R_a \subseteq W$ and $I_a \cap W = \emptyset$, i.e. all the reactants of $a$ are in $W$,
 while none of the inhibitors of $a$ are in $W$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
is defined by:
$\mathit{res}_a(W) \defeq P_a$, if $\mathit{en}_a(W)$, and $\mathit{res}_a(W) \defeq \emptyset$ otherwise.
\item Let $A$ be a finite set of reactions over $S$. The result of $A$ on $W$, denoted 
by $\mathit{res}_A(W)$,  is defined
by: $\mathit{res}_A(W) \defeq \bigcup_{a \in A} \mathit{res}_a(W)$.
\end{enumerate}
\end{definition}

>>>>

\begin{definition}[Reaction]\label{def:reaction}
Let $S$ be a set of entities.
A reaction over $S$ is a triple $a = (R,I,P)$, where $R, I, P$ are finite,  
non empty subsets of $S$ and
$R \cap I = \emptyset$. 
%If $S$ is a set such that  $R, I, P \subseteq S$, 
%then $a$ is a reaction in $S$.
\end{definition}

@@

\begin{definition}[Reaction Result]
Given a set of entities $S$, let $W\subseteq S$ be a finite subset of  entities.
\begin{enumerate}
\item Let $a\in \mathit{rac}(S)$ be a reaction over $S$. Then $a$ is enabled by the entities in the set $W$, denoted by $\mathit{en}_a(W)$, 
if $R_a \subseteq W$ and $I_a \cap W = \emptyset$, i.e. all the reactants of $a$ are in $W$,
 while none of the inhibitors of $a$ are in $W$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
is defined by:
$\mathit{res}_a(W) \defeq P_a$, if $\mathit{en}_a(W)$, and $\mathit{res}_a(W) \defeq \emptyset$ otherwise.
\item Let $A\subseteq \mathit{rac}(S)$ be a (finite) set of reactions over $S$. The result of $A$ on $W$, denoted 
by $\mathit{res}_A(W)$,  is defined
by: $\mathit{res}_A(W) \defeq \bigcup_{a \in A} \mathit{res}_a(W)$.
\end{enumerate}
\end{definition}

### 2020-08-14 TCS_v3/05-translation.tex
nuovi file TCS v3

 \begin{restatable}[]{lemma}{lemmastruct}
 \label{lem:struct}
%\begin{lemma}
 %\label{lem:struct}
  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~encoding.
  %{\color{red} embedding.}
  %translation. 
  If $exists$ $P'$ such that $P \xrightarrow{\restrict{\mathit{names}}\upsilon} P'$ is a transition of $P$, then
  \begin{enumerate}
  \item for each reaction  $aj \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s \in S$, the corresponding channel $s$ (suitably decorated) appear in $\upsilon$; 
  the channel $\mathit{cxt}$ appears in $\upsilon$;
  %processes $P_s$ (or $\overline{P_s}$) 
%  for each  $Cxt_s$ are
 %involved in the transition $t$;
  \item for each reaction $aj \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}_1$ is  overlapped by exactly one solid link offered by  processes representing entities.
  %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
  \end{enumerate}
  \end{restatable}

>>>>

 \begin{restatable}[]{lemma}{lemmastruct}
 \label{lem:struct}
%\begin{lemma}
 %\label{lem:struct}
  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~encoding.
  %{\color{red} embedding.}
  %translation. 
  If $exists$ $P'$ such that $P \xrightarrow{\restrict{\mathit{names}}\upsilon} P'$ is a transition of $P$, then
  \begin{enumerate}
  \item for each reaction  $aj \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s \in S$, the corresponding channel $s$ (suitably decorated) appear in $\upsilon$; 
  the channel $\mathit{cxt}$ appears in $\upsilon$;
  %processes $P_s$ (or $\overline{P_s}$) 
%  for each  $Cxt_s$ are
 %involved in the transition $t$;
  \item for each reaction $aj \in A$ and
  % each entity $s \in S$, 
  each virtual link offered by processes $P_a$ and $\mathit{Cxt}_1$ is  overlapped by exactly one solid link offered by  processes representing entities.
  %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
  \end{enumerate}
  \end{restatable}
